
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install libraries

```{r}
suppressPackageStartupMessages({
  library(DBI)
  library(RPostgres)
  library(dplyr)
  library(stringr)
  library(purrr)
  library(tidyr)
  library(jsonlite)
  library(stringi)
  library(readr)
  library(tibble)
})

```


# Get tiktok data from postgres

```{r}
# connect to postgres
con <- dbConnect(
  RPostgres::Postgres(),
  host = Sys.getenv("DB_HOST"),
  port = as.integer(Sys.getenv("DB_PORT", "5432")),
  dbname = Sys.getenv("DB_NAME"),
  user = Sys.getenv("DB_USER"),
  password = Sys.getenv("DB_PASS"),
  sslmode = "require"
)

# read in data
tiktok <- dbGetQuery(con, "SELECT * FROM clean_data.tik_tok")

```

# Helper functions to normalize text columns

```{r}
normalize_text <- function(x) {
  x <- dplyr::coalesce(as.character(x), "")
  x <- iconv(x, to = "UTF-8", sub = "")
  x <- stringi::stri_trans_general(x, "Latin-ASCII")
  x <- tolower(x)
  x <- str_replace_all(x, "[|,/;]+", " ")
  x <- str_replace_all(x, "#", " ")
  x <- str_replace_all(x, "@", " ")
  x <- str_replace_all(x, "[^a-z0-9 ]+", " ")
  x <- str_squish(x)
  x
}

normalize_compact <- function(x) {
  x <- normalize_text(x)
  str_replace_all(x, "\\s+", "")
}

clean_code <- function(x) {
  x <- as.character(x)
  x <- str_trim(x)
  x[x %in% c("", "\\N", "-", "N/A", "NA", "NULL", "null")] <- NA_character_
  x
}
```

# Blacklist / glue words

```{r}
bad_single_word_terms <- c(
  "nice", "reading", "allen", "tours", "lot", "can", "for", "hear",
  "same", "man", "part", "make", "date", "side", "tour", "travel",
  "terminal", "flight", "flights", "airport", "airline", "airlines",
  "airways", "city", "holiday", "vacation", "adventure", "sale",
  "surprise", "forest", "bend", "most", "normal", "guide", "hotel", "taylor"
)

glue_words <- c(
  "travel", "tourism", "tourist", "trip", "trips", "vacation", "vacations",
  "holiday", "holidays", "visit", "visiting", "guide", "guides",
  "adventure", "adventures", "explore", "exploring", "destination",
  "destinations", "getaway", "getaways", "journey", "journeys",
  "life", "food", "style", "crew", "sunset", "sunsets", "beauty",
  "beach", "summer", "winter", "check", "vibes", "living", "love",
  "diaries", "tips", "itinerary", "resort", "hotel"
)

glue_words <- unique(normalize_compact(glue_words))
```

# File readers

## alternate cities & us geo ref file

```{r}

read_alternate_names <- function(path) {
  readr::read_tsv(
    path,
    col_names = c(
      "alternateNameId", "geonameid", "isolanguage", "alternate_name",
      "isPreferredName", "isShortName", "isColloquial", "isHistoric",
      "from", "to"
    ),
    show_col_types = FALSE,
    quote = ""
  ) %>%
    transmute(
      geonameid = as.character(geonameid),
      isolanguage = as.character(isolanguage),
      alternate_name = as.character(alternate_name),
      isPreferredName = suppressWarnings(as.integer(isPreferredName)),
      isShortName = suppressWarnings(as.integer(isShortName)),
      isColloquial = suppressWarnings(as.integer(isColloquial)),
      isHistoric = suppressWarnings(as.integer(isHistoric))
    )
}

read_uscities <- function(path) {
  readr::read_csv(path, show_col_types = FALSE) %>%
    transmute(
      city = as.character(city),
      city_ascii = as.character(city_ascii),
      state_id = as.character(state_id),
      state_name = as.character(state_name),
      county_name = as.character(county_name),
      lat = suppressWarnings(as.numeric(lat)),
      lng = suppressWarnings(as.numeric(lng)),
      population = suppressWarnings(as.numeric(population)),
      timezone = as.character(timezone),
      uscity_id = as.character(id)
    )
}
```

# Country aliases
```{r}
build_country_aliases <- function(countries) {
  countries2 <- countries %>%
    mutate(
      country_name = as.character(Name),
      iso2 = clean_code(Code)
    )

  bind_rows(
    countries2 %>%
      transmute(
        match_type = "country",
        canonical_name = country_name,
        country_name = country_name,
        state_name = NA_character_,
        city_name = NA_character_,
        geonameid = NA_character_,
        city_key = NA_character_,
        state_key = normalize_compact(country_name),
        country_key = normalize_compact(country_name),
        alias = country_name
      ),
    tibble(
      match_type = "country",
      canonical_name = c("United States", "United Kingdom", "Germany"),
      country_name = c("United States", "United Kingdom", "Germany"),
      state_name = NA_character_,
      city_name = NA_character_,
      geonameid = NA_character_,
      city_key = NA_character_,
      state_key = c("unitedstates", "unitedkingdom", "germany"),
      country_key = c("unitedstates", "unitedkingdom", "germany"),
      alias = c("usa", "england", "deutschland")
    )
  ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    filter(alias_norm != "", !(alias_norm %in% bad_single_word_terms)) %>%
    distinct(alias_norm, canonical_name, .keep_all = TRUE)
}

```

# USA alias table

```{r}
build_us_state_aliases <- function(uscities) {
  state_ref <- uscities %>%
    distinct(state_name, state_id) %>%
    filter(!is.na(state_name), state_name != "")

  state_ref %>%
    transmute(
      match_type = "state",
      canonical_name = state_name,
      country_name = "United States",
      state_name = state_name,
      city_name = NA_character_,
      geonameid = NA_character_,
      city_key = NA_character_,
      state_key = normalize_compact(paste(state_name, "United States")),
      country_key = normalize_compact("United States"),
      alias = state_name
    ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    filter(alias_norm != "", !(alias_norm %in% bad_single_word_terms)) %>%
    distinct(alias_norm, canonical_name, .keep_all = TRUE)
}
```

## global city master file
```{r}
build_global_city_master <- function(cities, countries, alt_names,
                                     min_population = 5000) {

  countries_small <- countries %>%
    transmute(
      country_code = clean_code(Code),
      country_name = as.character(Name)
    )

  base <- cities %>%
    mutate(
      geonameid = as.character(geonameid),
      name = as.character(name),
      asciiname = as.character(asciiname),
      country_code = clean_code(country_code),
      admin1_code = clean_code(admin1_code),
      feature_code = as.character(feature_code),
      population = suppressWarnings(as.numeric(population)),
      latitude = suppressWarnings(as.numeric(latitude)),
      longitude = suppressWarnings(as.numeric(longitude))
    ) %>%
    filter(
      feature_code %in% c("PPLC", "PPLA", "PPLA2", "PPLA3", "PPLA4", "PPL") |
        (!is.na(population) & population >= min_population)
    ) %>%
    left_join(countries_small, by = "country_code") %>%
    filter(!is.na(country_name), country_name != "") %>%
    mutate(
      city_name = coalesce(asciiname, name),
      state_name = admin1_code,
      city_key = normalize_compact(paste(city_name, state_name, country_name)),
      state_key = normalize_compact(paste(state_name, country_name)),
      country_key = normalize_compact(country_name)
    ) %>%
    select(
      geonameid, city_name, state_name, country_name, country_code,
      latitude, longitude, population, city_key, state_key, country_key,
      name, asciiname
    )

  # Keep only alternate names for retained cities
  alt_city <- alt_names %>%
    filter(!is.na(alternate_name), alternate_name != "") %>%
    semi_join(base %>% select(geonameid), by = "geonameid") %>%
    filter(is.na(isHistoric) | isHistoric != 1L) %>%
    filter(nchar(alternate_name) >= 2) %>%
    transmute(geonameid, alias = alternate_name)

  base_aliases <- bind_rows(
    base %>% transmute(geonameid, alias = name),
    base %>% transmute(geonameid, alias = asciiname),
    alt_city
  ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    filter(
      alias_norm != "",
      !(alias_norm %in% bad_single_word_terms),
      nchar(alias_norm) >= 2
    ) %>%
    distinct(geonameid, alias_norm, .keep_all = TRUE)

  base %>%
    inner_join(base_aliases, by = "geonameid")
}

```

## USA city master file
```{r}
build_us_city_master <- function(uscities, min_population = 2000) {
  bind_rows(
    uscities %>%
      mutate(population = suppressWarnings(as.numeric(population))) %>%
      filter(!is.na(city_ascii), city_ascii != "") %>%
      filter(is.na(population) | population >= min_population) %>%
      transmute(
        geonameid = paste0("uscity_", uscity_id),
        city_name = city_ascii,
        state_name = state_name,
        country_name = "United States",
        country_code = "US",
        latitude = lat,
        longitude = lng,
        population = population,
        city_key = normalize_compact(paste(city_ascii, state_name, "United States")),
        state_key = normalize_compact(paste(state_name, "United States")),
        country_key = normalize_compact("United States"),
        alias = city
      ),
    uscities %>%
      mutate(population = suppressWarnings(as.numeric(population))) %>%
      filter(!is.na(city_ascii), city_ascii != "") %>%
      filter(is.na(population) | population >= min_population) %>%
      transmute(
        geonameid = paste0("uscity_", uscity_id),
        city_name = city_ascii,
        state_name = state_name,
        country_name = "United States",
        country_code = "US",
        latitude = lat,
        longitude = lng,
        population = population,
        city_key = normalize_compact(paste(city_ascii, state_name, "United States")),
        state_key = normalize_compact(paste(state_name, "United States")),
        country_key = normalize_compact("United States"),
        alias = city_ascii
      )
  ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    filter(alias_norm != "", !(alias_norm %in% bad_single_word_terms)) %>%
    distinct(geonameid, alias_norm, .keep_all = TRUE)
}

```

## Airport alias (assigned to city, state, country for rollup)
```{r}
build_airport_aliases <- function(airports, uscities) {
  airports_small <- airports %>%
    transmute(
      airport_id = as.character(id),
      airport_name = as.character(name),
      iata_code = clean_code(iata_code),
      gps_code = clean_code(gps_code),
      ident = clean_code(ident),
      municipality = as.character(municipality),
      iso_country = clean_code(iso_country),
      iso_region = clean_code(iso_region)
    )

  us_lookup <- uscities %>%
    transmute(
      city_norm = normalize_text(city_ascii),
      state_id = as.character(state_id),
      state_name = as.character(state_name),
      city_name = as.character(city_ascii),
      country_name = "United States",
      city_key = normalize_compact(paste(city_ascii, state_name, "United States")),
      state_key = normalize_compact(paste(state_name, "United States")),
      country_key = normalize_compact("United States")
    ) %>%
    distinct(city_norm, state_id, .keep_all = TRUE)

  airport_city_xwalk <- airports_small %>%
    mutate(
      city_norm = normalize_text(municipality),
      state_id = if_else(str_detect(iso_region, "-"),
                         str_replace(iso_region, ".*-", ""),
                         iso_region)
    ) %>%
    left_join(us_lookup, by = c("city_norm", "state_id")) %>%
    filter(!is.na(city_key))

  bind_rows(
    airport_city_xwalk %>%
      transmute(
        match_type = "airport",
        canonical_name = airport_name,
        country_name, state_name, city_name,
        geonameid = airport_id,
        city_key, state_key, country_key,
        alias = airport_name
      ),
    airport_city_xwalk %>%
      filter(!is.na(iata_code)) %>%
      transmute(
        match_type = "airport_code",
        canonical_name = airport_name,
        country_name, state_name, city_name,
        geonameid = airport_id,
        city_key, state_key, country_key,
        alias = iata_code
      ),
    airport_city_xwalk %>%
      filter(!is.na(gps_code)) %>%
      transmute(
        match_type = "airport_code",
        canonical_name = airport_name,
        country_name, state_name, city_name,
        geonameid = airport_id,
        city_key, state_key, country_key,
        alias = gps_code
      ),
    airport_city_xwalk %>%
      filter(!is.na(ident)) %>%
      transmute(
        match_type = "airport_code",
        canonical_name = airport_name,
        country_name, state_name, city_name,
        geonameid = airport_id,
        city_key, state_key, country_key,
        alias = ident
      )
  ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    filter(alias_norm != "") %>%
    distinct(match_type, alias_norm, city_key, .keep_all = TRUE)
}

```

# All alias into one lookup
```{r}

build_place_alias_table <- function(
  airports, cities, countries, alt_names, uscities,
  min_global_city_population = 5000,
  min_us_city_population = 2000
) {
  global_city_master <- build_global_city_master(
    cities = cities,
    countries = countries,
    alt_names = alt_names,
    min_population = min_global_city_population
  )

  us_city_master <- build_us_city_master(
    uscities = uscities,
    min_population = min_us_city_population
  )

  city_aliases <- bind_rows(
    global_city_master %>%
      transmute(
        match_type = "city",
        canonical_name = city_name,
        country_name, state_name, city_name,
        geonameid, city_key, state_key, country_key,
        alias = alias_norm
      ),
    us_city_master %>%
      transmute(
        match_type = "city",
        canonical_name = city_name,
        country_name, state_name, city_name,
        geonameid, city_key, state_key, country_key,
        alias = alias_norm
      )
  ) %>%
    mutate(
      alias_norm = normalize_text(alias),
      alias_compact = normalize_compact(alias)
    ) %>%
    distinct(alias_norm, city_key, .keep_all = TRUE)

  state_aliases <- build_us_state_aliases(uscities)
  country_aliases <- build_country_aliases(countries)
  airport_aliases <- build_airport_aliases(airports, uscities)

  alias_tbl <- bind_rows(
    city_aliases,
    state_aliases,
    country_aliases,
    airport_aliases
  ) %>%
    mutate(
      n_words = str_count(alias_norm, "\\S+") + if_else(alias_norm == "", 0L, 0L)
    ) %>%
    filter(alias_norm != "", alias_compact != "") %>%
    distinct(match_type, alias_norm, city_key, state_key, country_key, .keep_all = TRUE)

  alias_tbl
}
```

## Fast hash lookup
```{r}
build_geo_lookup_fast <- function(alias_tbl, max_phrase_words = 6L) {
  alias_tbl <- alias_tbl %>%
    mutate(
      n_words = str_count(alias_norm, "\\S+") + 1L
    )

  exact_lookup <- split(alias_tbl, alias_tbl$alias_norm)
  compact_lookup <- split(alias_tbl, alias_tbl$alias_compact)

  phrase_sizes <- sort(unique(alias_tbl$n_words))
  phrase_sizes <- phrase_sizes[phrase_sizes >= 1L & phrase_sizes <= max_phrase_words]

  # named logical vectors for O(1)-ish membership tests
  exact_keys <- stats::setNames(rep(TRUE, length(exact_lookup)), names(exact_lookup))
  compact_keys <- stats::setNames(rep(TRUE, length(compact_lookup)), names(compact_lookup))

  list(
    exact_lookup = exact_lookup,
    compact_lookup = compact_lookup,
    exact_keys = exact_keys,
    compact_keys = compact_keys,
    phrase_sizes = phrase_sizes,
    glue_words = glue_words
  )
}

```

Help function for phrase generation:
```{r}
make_ngrams <- function(tokens, k) {
  n <- length(tokens)
  if (n < k) return(character(0))
  if (k == 1L) return(tokens)

  starts <- seq_len(n - k + 1L)
  vapply(
    starts,
    function(i) paste(tokens[i:(i + k - 1L)], collapse = " "),
    character(1)
  )
}
```

# matcher
```{r}

extract_geo_tags_one_fast <- function(text_vec, lookup) {
  text_vec <- text_vec[!is.na(text_vec)]
  text_vec <- as.character(text_vec)
  text_vec <- text_vec[nzchar(text_vec)]

  if (length(text_vec) == 0) return("[]")

  text_norm <- normalize_text(paste(text_vec, collapse = " "))
  if (!nzchar(text_norm)) return("[]")

  tokens <- unlist(str_split(text_norm, "\\s+"), use.names = FALSE)
  tokens <- tokens[nzchar(tokens)]
  if (length(tokens) == 0) return("[]")

  hits <- vector("list", 0L)

  # 1) exact phrase match with hashed membership checks
  for (k in lookup$phrase_sizes) {
    grams <- unique(make_ngrams(tokens, k))
    if (length(grams) == 0) next

    keep <- !is.na(lookup$exact_keys[grams])
    if (any(keep)) {
      matched <- grams[keep]
      hits <- c(hits, lookup$exact_lookup[matched])
    }
  }

  # 2) compact token matches like newyork / usa / atl
  token_compact <- unique(normalize_compact(tokens))
  token_compact <- token_compact[nzchar(token_compact)]

  if (length(token_compact) > 0) {
    keep_compact <- !is.na(lookup$compact_keys[token_compact])
    if (any(keep_compact)) {
      matched_compact <- token_compact[keep_compact]
      hits <- c(hits, lookup$compact_lookup[matched_compact])
    }
  }

  # 3) prefix/suffix glue stripping only
  # catches sandiegolife, stockholmsunset, easyjetcrew, etc.
  if (length(token_compact) > 0) {
    stripped <- character(0)

    for (tok in token_compact) {
      tok_n <- nchar(tok)
      for (glue in lookup$glue_words) {
        glue_n <- nchar(glue)
        if (tok_n <= glue_n) next

        if (startsWith(tok, glue)) {
          stripped <- c(stripped, substr(tok, glue_n + 1L, tok_n))
        }
        if (endsWith(tok, glue)) {
          stripped <- c(stripped, substr(tok, 1L, tok_n - glue_n))
        }
      }
    }

    stripped <- unique(stripped[nzchar(stripped)])
    if (length(stripped) > 0) {
      keep_stripped <- !is.na(lookup$compact_keys[stripped])
      if (any(keep_stripped)) {
        matched_stripped <- stripped[keep_stripped]
        hits <- c(hits, lookup$compact_lookup[matched_stripped])
      }
    }
  }

  if (length(hits) == 0) return("[]")

  out <- bind_rows(hits) %>%
    distinct(
      match_type, canonical_name, country_name, state_name, city_name,
      city_key, state_key, country_key, alias_norm,
      .keep_all = TRUE
    ) %>%
    transmute(
      matched_as = match_type,
      matched_text = alias_norm,
      canonical_name = canonical_name,
      city = city_name,
      state = state_name,
      country = country_name,
      city_key = city_key,
      state_key = state_key,
      country_key = country_key
    ) %>%
    arrange(country, state, city, canonical_name)

  jsonlite::toJSON(out, auto_unbox = TRUE)
}


extract_geo_tags_df_fast <- function(df, lookup, text_cols, out_col = "geo_tag") {
  stopifnot(all(text_cols %in% names(df)))

  text_mat <- df[, text_cols, drop = FALSE]

  df[[out_col]] <- vapply(
    seq_len(nrow(text_mat)),
    function(i) extract_geo_tags_one_fast(unlist(text_mat[i, ], use.names = FALSE), lookup),
    character(1)
  )

  df
}
```

# Config

```{r}
# geo files
airports  <- read.csv("geo_datasets/airports.csv", stringsAsFactors = FALSE)
airlines  <- read.csv("geo_datasets/airlines.csv", stringsAsFactors = FALSE)
cities    <- read.csv("geo_datasets/cities.csv", stringsAsFactors = FALSE)
countries <- read.csv("geo_datasets/countries.csv", stringsAsFactors = FALSE)
alt_names <- read_alternate_names("geo_datasets/alternatenames.txt")
uscities  <- read_uscities("geo_datasets/uscities.csv")

```

## Build alias table and lookup
```{r}
system.time({
  alias_tbl <- build_place_alias_table(
    airports = airports,
    cities = cities,
    countries = countries,
    alt_names = alt_names,
    uscities = uscities,
    min_global_city_population = 10000,
    min_us_city_population = 10000
  )
})

system.time({
  geo_lookup <- build_geo_lookup_fast(alias_tbl, max_phrase_words = 6L)
})
```

```{r}
# parameters
TEXT_COLS <- c("title","caption","description", "hashtags","channel")

```

# run
```{r}
system.time({
  tiktok <- extract_geo_tags_df_fast(
    df = tiktok,
    lookup = geo_lookup,
    text_cols = TEXT_COLS,
    out_col = "geo_tag"
  )
})
```


# Push data back to postgres

```{r mongo-query}
con <- dbConnect(
  RPostgres::Postgres(),
  host = Sys.getenv("DB_HOST"),
  port = as.integer(Sys.getenv("DB_PORT", "5432")),
  dbname = Sys.getenv("DB_NAME"),
  user = Sys.getenv("DB_USER"),
  password = Sys.getenv("DB_PASS"),
  sslmode = "require" 
)

# sanity check
dbGetQuery(con, "select now() as now_utc;")
```

``` {r}
schema_name <- "clean_data"
table_name  <- "tik_tok"    


# OVERWRITE table
dbWriteTable(
  con,
  Id(schema = schema_name, table = table_name),
  tiktok,
  overwrite = TRUE
)

# verify
dbGetQuery(con, sprintf("SELECT COUNT(*) AS n FROM %s.%s", schema_name, table_name))
```

```{r}
# close connection
dbDisconnect(con)
```